---
title: "Ecosystem Science and Sustainability 330"
author:
  - name: Libby Prince
format: html
subtitle: 'Quantitative Reasoning for Ecosystem Science'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## COVID-19 Data

### Data

> "We are providing two sets of data with cumulative counts of coronavirus cases and deaths: one with our most current numbers for each geography and another with historical data showing the tally for each day for each geography ... the historical files are the final counts at the end of each day ... The historical and live data are released in three files, one for each of these geographic levels: U.S., states and counties. A smaller file with only the most recent 30 days of data is also available"

```{r, eval = FALSE}
https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties-recent.csv
```

### Start

I will get you started this week, in the following code chunk, I am attaching the tidyverse package; saving the NY-Times URL as a value called "url"; and I am reading that URL into an object called `covid`

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties-recent.csv'
covid = read_csv(url)
head(covid, 5)
```

```{r, eval = FALSE}
slice(covid, 1:5)
```


## Question 1: 
The top five counties with the most current cases are Los Angeles, New York City, Miami-Dade, Cook, and Maricopa.

```{r}
covid %>%
  filter(date == max(date)) %>%
  arrange(desc(cases)) %>% 
  slice(1:5) 
```
## Question 2: 
The top states with the most current cases are California, Texas, Florida, New York, and Illinois.
 

```{r}
covid %>%
  mutate(date = as.Date(date)) %>%
  filter(date == max(date)) %>% 
  group_by(state) %>% 
  summarize(total_cases = sum(cases, na.rm = TRUE)) %>%
  arrange(desc(total_cases)) %>%
  slice(1:5)
 
 
```

## Question 3: 
The five counties with the worst current death/cases ratio are Storey, Sabine, McMullen, Blaine, and Grant.

Use the `dplyr` verbs to report the **5** counties with the _worst_ current death/cases ratio: (e.g.$100* (deaths/cases)$)


```{r}
covid %>%
  mutate(date = as.Date(date)) %>%
  filter(date == max(date), cases > 0, county != "Unknown") %>%
  mutate(death_case_ratio = 100 * deaths / cases) %>%
  arrange(desc(death_case_ratio)) %>%
  slice(1:5)      
```

## Question 4: 

The top five states with the worst current death/cases ratio are Pennsylvania, Michigan, Georgia, Nevada, and Arizona. 

```{r}
covid %>%
  mutate(date = as.Date(date)) %>%
  filter(date == max(date), cases > 0) %>%
  group_by(state) %>%                     
  summarize(total_cases = sum(cases, na.rm = TRUE),
            total_deaths = sum(deaths, na.rm = TRUE)) %>%
  mutate(death_case_ratio = 100 * total_deaths / total_cases) %>%  # 
  arrange(desc(death_case_ratio)) %>%     
  slice(1:5)                        

```
